Casual Bandits

### 2016-01-13 causal-bandit/icml-2015/lower-bound.tex
added some todos
incorporated the simple-regret upper bound and continued to cleanup the proof
updates

\newcommand{\simpleregret}{R_T^{\text{simple}}}

We prove a lower bound on the simple regret that matches up to logarithmic factors the upper bound given in \cref{thm:simple}

\begin{theorem}\label{thm:lower}
Suppose $\eta_t \sim \mathcal N(0, 1)$ and $\boldsymbol{q}$ satisfies $m(\boldsymbol{q}) = m$.
Then for all strategies there exists a reward function such that
\eq{
\simpleregret \geq \frac{\frac{m}{2e} - 1}{m} \sqrt{\frac{2m}{3T}} = \Omega\left(\sqrt{\frac{m}{T}}\right)\,.
}
\end{theorem}

\begin{proof}
Assume without loss of generality that $q_i \leq 1/2$ for all $i$ and that $q_1 \leq q_2 \leq \ldots \leq q_N$.
For each $i$ define reward function $r_i$ by
\eq{
r_0(\boldsymbol{X}) &= 0 &
r_i(\boldsymbol{X}) &= \begin{cases}
\epsilon & \text{if } X_i = 1 \\
0 & \text{otherwise}\,,
\end{cases}
}
where $\epsilon > 0$ is some constant to be chosen later.
We abbreviate $R_{T,i}^{\text{simple}}$ to be the expected simple regret incurred when interacting with the
environment determined by $\boldsymbol{q}$ and $r_i$. Let $P_i$ be the corresponding measure
on all observations over all $T$ rounds and $\mathbb E_i$ the expectation with respect to $P_i$. By Lemma 2.6 by \cite{Tsy08} we have
\eq{
\Prz{I_t = i} + \Pri{I_t \neq i} \geq \exp\left(-\KL(P_0, P_i)\right)\,,
}
where $\KL(P_0, P_i)$ is the KL divergence between measures $P_0$ and $P_i$.
Let $T_i(T) = \sum_{t=1}^T \ind{I_t = i}$ be the total number of times the learner intervenes on variable $i$ (with any intervention). 
Then for $i \leq m$ a computation shows that 
\eq{
\KL(P_0, P_i) 
&= \frac{\epsilon^2}{2} \mathbb E_0\left[\sum_{t=1}^T \ind{X_{t,i} = 1}\right] \\
&\leq \frac{\epsilon^2}{2} \left(\mathbb E_0 T_i(T) + q_i T\right) \\
&\leq \frac{\epsilon^2}{2} \left(\mathbb E_0 T_i(T) + \frac{T}{m}\right)\,.
}
Define set $A$ by
\eq{
A = \set{i \leq m : \mathbb E_0 T_i(T) \leq 2T / m}\,.
}
Then for $i \in A$ and choosing $\epsilon = \sqrt{2m/3T}$ we have
\eq{
\KL(P_0, P_i) \leq \frac{3T\epsilon^2}{2m} = 1\,. 
}
Now $\sum_{i=1}^m \mathbb E_0 T_i(T) \leq T$, which implies that $|A| \geq m/2$.
Therefore
\eq{
\sum_{i \in A} \Pri{I_t \neq i} 
&\geq \sum_{i \in A} \exp\left(-\KL(P_0, P_i)\right) - 1\\
&\geq \frac{|A|}{e} - 1 
\geq \frac{m}{2e} - 1\,.
}
Therefore there exists an $i \in A$ such that
\eq{
\Pri{I_t \neq i} \geq \frac{\frac{m}{2e} - 1}{m}\,.
}
Therefore
\eq{
R^{\text{simple}}_{T,i} \geq \frac{1}{2} P_i(I_t \neq i) \epsilon \geq \frac{\frac{m}{2e} - 1}{2m} \sqrt{\frac{2m}{3T}}
}
as required.
\end{proof}

>>>>

We prove a lower bound on the simple regret that matches up to logarithmic factors the upper bound given in Theorem \ref{thm:uq-simple}

\todot{Remark that the Gaussian assumption is convenient but that other noise models will also work. Eg., Bernoulli.}
\begin{theorem}\label{thm:lower}
Suppose $\eta_t \sim \mathcal N(0, 1)$ and $\boldsymbol{q}$ satisfies $m(\boldsymbol{q}) = m$.
Then for all strategies there exists a reward function such that
\eq{
\simpleregret \geq \frac{\frac{m}{2e} - 1}{m} \sqrt{\frac{2m}{3T}} = \Omega\left(\sqrt{\frac{m}{T}}\right)\,.
}
\end{theorem}

\begin{proof}
Assume without loss of generality that $q_i \leq 1/2$ for all $i$ and that $q_1 \leq q_2 \leq \ldots \leq q_N$.
For each $i$ define reward function $r_i$ by
\eq{
r_0(\boldsymbol{X}) &= 0 &
r_i(\boldsymbol{X}) &= \begin{cases}
\epsilon & \text{if } X_i = 1 \\
0 & \text{otherwise}\,,
\end{cases}
}
where $\epsilon > 0$ is some constant to be chosen later.
We abbreviate $R_{T,i}^{\text{simple}}$ to be the expected simple regret incurred when interacting with the
environment determined by $\boldsymbol{q}$ and $r_i$. Let $P_i$ be the corresponding measure
on all observations over all $T$ rounds and $\mathbb E_i$ the expectation with respect to $P_i$. By Lemma 2.6 by \cite{Tsy08} we have
\eq{
\Prz{I_t = i} + \Pri{I_t \neq i} \geq \exp\left(-\KL(P_0, P_i)\right)\,,
}
where $\KL(P_0, P_i)$ is the KL divergence between measures $P_0$ and $P_i$.
Let $T_i(T) = \sum_{t=1}^T \ind{I_t = i}$ be the total number of times the learner intervenes on variable $i$ (with any intervention). 
Then for $i \leq m$ a computation shows that 
\todot{Need to add at minimum some citations here}
\eq{
\KL(P_0, P_i) 
&= \frac{\epsilon^2}{2} \mathbb E_0\left[\sum_{t=1}^T \ind{X_{t,i} = 1}\right] \\
&\leq \frac{\epsilon^2}{2} \left(\mathbb E_0 T_i(T) + q_i T\right) \\
&\leq \frac{\epsilon^2}{2} \left(\mathbb E_0 T_i(T) + \frac{T}{m}\right)\,.
}
Define set $A$ by
\eq{
A = \set{i \leq m : \mathbb E_0 T_i(T) \leq 2T / m}\,.
}
Then for $i \in A$ and choosing $\epsilon = \sqrt{2m/3T}$ we have
\eq{
\KL(P_0, P_i) \leq \frac{3T\epsilon^2}{2m} = 1\,. 
}
Now $\sum_{i=1}^m \mathbb E_0 T_i(T) \leq T$, which implies that $|A| \geq m/2$.
Therefore
\eq{
\sum_{i \in A} \Pri{I_t \neq i} 
&\geq \sum_{i \in A} \exp\left(-\KL(P_0, P_i)\right) - 1\\
&\geq \frac{|A|}{e} - 1 
\geq \frac{m}{2e} - 1\,.
}
Therefore there exists an $i \in A$ such that
\eq{
\Pri{I_t \neq i} \geq \frac{\frac{m}{2e} - 1}{m}\,.
}
Therefore
\eq{
R^{\text{simple}}_{T,i} \geq \frac{1}{2} P_i(I_t \neq i) \epsilon \geq \frac{\frac{m}{2e} - 1}{2m} \sqrt{\frac{2m}{3T}}
}
as required.
\end{proof}

### 2016-01-13 causal-bandit/icml-2015/setup.tex
updates

The game proceeds over $T$ identical rounds (or time-steps).
In each round $t$ the learner can choose either to do nothing or they can choose a variable $I_t \in \set{1,\ldots,N}$ and
an intervention $J_t \in \set{0,1}$. After the learner has chosen an intervention they observe $X_{t,i} \in \set{0,1}$ for all $i$ where
\eq{
X_{t,i} \sim \begin{cases}
\dirac(J_t) & \text{if } I_t = i \\
\bernoulli(q_i) & \text{otherwise}\,.
\end{cases}
}
where $\boldsymbol{q} \in [0,1]^N$ is a (possibly unknown) vector of probabilities with $q_i = \P{X_i = 1}$. Note that if the learner did not choose an intervention then $I_t$ is undefined
and $X_{t,i} \sim \bernoulli(q_i)$ for all variables $i \in \set{1,\ldots,N}$.
Finally the learner observes the reward $Y_t = r(X_t) + \eta_t$ where 
\eq{
r : \set{0,1}^N \to \R
}
is arbitrary (and unknown) 
and $\eta_t$ is a $1$-subgaussian noise
term (with the distribution possibly dependent on $X_t$). 
\todom{Connect to opening ex.}

@@

We consider two standard performance measures. The first is the cumulative regret, which measures the difference between the reward expected under
the omnipotent strategy that knows the optimal action in advance and the expected reward of the learner. 
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
The second performance measure is the simple regret, which measures the performance of the learner on the final round only.
\eqn{
\label{eq:regret-simple}
r_T = \mu^* - \E{\mu_{I_T,J_T}}\,.
}
Both versions of the regret are useful in different circumstances. The first is most useful when the learner is truly online and can change their policy over time.
The second is useful when the exploration budget is limited and a fixed policy should eventually be chosen. \todot{make this nicer}

>>>>

The game proceeds over $T$ identical rounds (or time-steps).
In each round $t$ the learner can choose either to do nothing (denoted by $I_t = \bot$) or they can choose $I_t \in \set{1,\ldots,N}$ and
an intervention $J_t \in \set{0,1}$. After the learner has chosen an intervention they observe $X_{t,i} \in \set{0,1}$ for all $i$ where
\eq{
X_{t,i} \sim \begin{cases}
\dirac(J_t) & \text{if } I_t = i \\
\bernoulli(q_i) & \text{otherwise}\,.
\end{cases}
}
where $\boldsymbol{q} \in [0,1]^N$ is a (possibly unknown) vector of probabilities with $q_i = \P{X_i = 1}$. 
Note that if the learner did not choose an intervention ($I_t =  \bot$), then
$X_{t,i} \sim \bernoulli(q_i)$ for all variables $i \in \set{1,\ldots,N}$.
Finally the learner observes the reward $Y_t = r(X_t) + \eta_t$ where 
\eq{
r : \set{0,1}^N \to \R
}
is arbitrary (and unknown) 
and $\eta_t$ is a $1$-subgaussian noise
term (with the distribution possibly dependent on $X_t$). 
In the farming example $X_{t,1}$ might represent whether the plants had sufficient water in the $t$th season (controllable by installing irrigation) 
while $X_{t,2}$ might be whether or not there was a spring frost (controllable with heat-lamps or covers).

@@

We consider two standard performance measures. The first is the cumulative regret, which measures the difference between the reward expected under
the omnipotent strategy that knows the optimal action in advance and the expected reward of the learner. 
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
The second performance measure is the simple regret, which measures the performance of the learner on the final round only.
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{I_T,J_T}}\,.
}
The simple regret typically makes more sense when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. The cumulative
regret is more appropriate for online agents that may continue to adapt their policy.
